---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# pickleR

<!-- badges: start -->
[![R-CMD-check](https://github.com/d-sharpe/pickleR/workflows/R-CMD-check/badge.svg)](https://github.com/d-sharpe/pickleR/actions)
[![test-coverage](https://codecov.io/gh/d-sharpe/pickleR/branch/master/graph/badge.svg?token=5WSPCOJ0EF)](https://codecov.io/gh/d-sharpe/pickleR)
<!-- badges: end -->

*pickleR* provides for the serializing and de-serializing of an R object structure,
similar to that seen in the Python module [pickle](https://docs.python.org/3/library/pickle.html).
*pickleR* traverses the object hierarchy, keeping track of any sub-objects it has already encountered
so that they will not be stored again. Therefore, recursive, nested and shared objects are handled by *pickleR*.

The unpickling process reconstitutes the object with shared objects and recursions intact.


## Installation

You can install the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("d-sharpe/pickleR")
```

## Example

```{r example}
library(pickleR)
library(lobstr)

valueVector <- rnorm(1000)

## memory address of vector
obj_addr(valueVector)

## object memory size
obj_size(valueVector)

listOfValues <-
  list(valueVector,
       valueVector)

## Each element of the listOfValues points to the same memory address of the
## original valueVector

lobstr::sxp(x = listOfValues)

## listOfValues memory size
obj_size(listOfValues)

RListOfValues <-
  unserialize(serialize(listOfValues, connection = NULL))

## Each element of the RListOfValues points to a new vector i.e. shared object
## reference were not maintained

lobstr::sxp(x = RListOfValues)

## RListOfValues memory size
obj_size(RListOfValues)

# The serialized object consumes more memory than the original

pickleListOfValues <-
  unpickle(pickle(listOfValues))

## Each element of the pickleListOfValues points to the same memory 
## address i.e. shared object

lobstr::sxp(x = pickleListOfValues)

## pickleListOfValues memory size
obj_size(pickleListOfValues)

# The pickled object consumes the same memory as the original
```
