---
title: "Using pickleR to (de-)serialize while maintaining object references"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using pickleR to (de-)serialize while maintaining object references}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  eval = TRUE,
  cache = FALSE,
  comment = "#>"
)
```

*pickleR* is an alternative to the in-built `serialize` and `unserialize` functions. When an object is serialized and subsequently deserialized using these functions, the intra-object references to the same element are usually not wholly perserved (weak references may be maintained). The most obvious effect of this for simple objects (e.g. lists) is an increased object memory footprint of the restored object compared to the original.

## Built-in serialize of simple objects

Consider the following example:

```{r}
integerVector <- runif(10000)

listOfIntegerVectors <- list(integerVector, integerVector, integerVector)

library(lobstr)

obj_addr(integerVector)

sxp(listOfIntegerVectors)
```

Each element of `listOfIntegerVectors` points to the same memory address, which is that of `integerVector`.

And we can see that the memory footprint of the `listOfIntegerVectors` is the almost the same as the `integerVector`. In this case 80 Bytes is consumed by the list structure.

```{r}
obj_size(integerVector)

obj_size(listOfIntegerVectors)
```

When the `listOfIntegerVectors` is serialized to a raw vector and restored, the elements of the restored object no longer share a common memory address. Also note that the raw vector is approximately 3 times the size of `integerVector` indicating that 3 copies of the vector have been saved.

```{r}
serializedListOfIntegerVectors <-
  serialize(listOfIntegerVectors, connection = NULL)

obj_size(serializedListOfIntegerVectors)

restoredListOfIntegerVectors <-
  unserialize(serializedListOfIntegerVectors)

sxp(restoredListOfIntegerVectors)
```

But now the list contains three copies of `integerVector` in memory for the `restoredListOfIntegerVectors`. We can see that the restored list object consumes increased memory compared to the original.

```{r}
obj_size(integerVector)

obj_size(listOfIntegerVectors)
# Bytes of integerVector + 80 B list structure

obj_size(restoredListOfIntegerVectors)
# 3 * bytes of integerVector + 80 B list structure
```

## PickleR serialize of simple objects

When the `listOfIntegerVectors` is _pickled_ to a raw vector and restored, the elements of the restored object continue to share a common memory address. Also note that the raw vector is of approximately equal size of `integerVector` indicating that only one copy of the vector have been saved.

```{r}
library(pickleR)

pickledListOfIntegerVectors <-
  pickle(listOfIntegerVectors, connection = NULL)

obj_size(pickledListOfIntegerVectors)

unpickledListOfIntegerVectors <-
  unpickle(pickledListOfIntegerVectors)

sxp(unpickledListOfIntegerVectors)
```

But now the list contains three copies of `integerVector` in memory for the `unpickledListOfIntegerVectors`. We can see that the restored list object consumes increased memory compared to the original.

```{r}
obj_size(integerVector)

obj_size(listOfIntegerVectors)
# Bytes of integerVector + 80 B list structure

obj_size(unpickledListOfIntegerVectors)
# Bytes of integerVector + 80 B list structure
```

## Serialization of complex objects

The loss of shared memory addresses in simple objects is not really a big problem. If a user alters a value in the `integerVector` R's copy-on-modify semantics would create a new vector with a new memory address and other 
elements would not receive that alteration anyway.

```{r}
listOfIntegerVectors[[1]][1] <- 5L

listOfIntegerVectors[[1]][1]

listOfIntegerVectors[[2]][1]

listOfIntegerVectors[[3]][1]

sxp(listOfIntegerVectors)
```

A more interesting problem is seen with nested, recursive and self-referencing environments as can be created in the `R6` package.

Suppose we have `R6` classes where circular referencing exists:

```{r}
library(R6)

child <-
  R6Class(
    classname = "child",
    class = TRUE,
    cloneable = TRUE,
    public = list(
      parent = NULL,
      initialize = function(parent = NULL, ...) {
        self$parent = parent
      },
      normalValue = rnorm(1),
                  uniformValue = runif(1),
                  normalRange = function(mean, sd, n = 10) {
                    return(rnorm(n = n, mean = mean, sd = sd))
                  },
                  uniformRange = function(min, max, n = 10) {
                    return(runif(n = n, min = min, max = max))
                  })
  )

parent <- 
  R6Class(
    classname = "parent",
    class = TRUE,
    cloneable = TRUE,
    public = list(
      children = list(),
      value = 1,
      initialize = function(...) {
        self$children <- lapply(1:5, function(i) {
          return(child$new(parent = self))
        })
      }
    )
  )
```

In this case the `parent` class has 5 `child` instances stored in a list (_has-a_ association). Each child class contains a reference to it the `parent` class that created it.

If we create an instance of the parent class and examine it:

```{r}
instance1 <- parent$new()

obj_addr(instance1)

obj_size(instance1)
```

And then extract and examine the `parent` object stored within each `child`

```{r}
obj_addrs(lapply(instance1$children, function(child) {child$parent}))
```

Each child correctly references the same instance of the `parent` class.

If we examine the body of a function in each child, we can see that they all reference the same object (from the class generator).

```{r}
obj_addrs(lapply(instance1$children, function(child) {
  body(child$normalRange)
}))
```

`R6` classes can be serialized and unserialized using the built-in methods, but the references are not preserved. Note that the restored `R6` object is approximately 4.3 times the size of the original.

```{r}
serializedR6Object <- 
  serialize(instance1, connection = NULL)

obj_size(serializedR6Object)

restoredR6Object <-
  unserialize(serializedR6Object)

obj_size(restoredR6Object)

obj_addrs(lapply(restoredR6Object$children, function(child) {child$parent}))
```

In the restored `R6` object the function bodies are not inherited from the generator and point to different memory addresses.

```{r}
obj_addrs(lapply(restoredR6Object$children, function(child) {
  body(child$normalRange)
}))
```

With _pickleR_ this not the case. The function bodies of each child are shared references to a single object, however, this object is not the same as that from the generator. Of note, that the raw value of the pickled object is approximately half that of the serialized object, and the unpickled object is smaller than the original. The unpickled object is smaller because references to the source files for function definitions are stripped from function bodies in pickleR. This does not impact the objects functionality, but when viewing the function definition comments, etc will be lost.

```{r}
pickledR6Object <- 
  pickle(instance1, connection = NULL)

obj_size(pickledR6Object)

unpickleddR6Object <-
  unpickle(pickledR6Object)

obj_size(unpickleddR6Object)

obj_addrs(lapply(unpickleddR6Object$children, function(child) {child$parent}))

obj_addrs(lapply(unpickleddR6Object$children, function(child) {
  body(child$normalRange)
}))
```

In significantly more complex systems with higher levels of class composition and recursion the effect of _pickleR_ vs the in-built `serialize` and `unserialize` can be dramatic.
