---
title: "Using pickleR to (de-)serialize while maintaining object references"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Using pickleR to (de-)serialize while maintaining object references}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(pickleR)
```

Consider the following example:

```{r}
x <- 1:100

list_of_x <- list(x, x, x)
```

Each element of `list_of_x` points to the same memory address of `x`.

```{r}
library(pryr)

pryr::inspect(x)

pryr::inspect(list_of_x)
```

However, when the `list_of_x` object is `serialized()` to a raw vector and restored via
`unserialize()`, the elements of the stored object \strong{do not} share a common target
as indicated by the different memory address of each element.

```{r}
restored_list_of_x <-
  unserialize(serialize(list_of_x, connection = NULL))

pryr::inspect(restored_list_of_x)

identical(list_of_x, restored_list_of_x)
```

Therefore, we now have three copies of `x` in memory for the `restored_list_of_x`.

```{r}
pryr::object_size(x)

pryr::object_size(list_of_x)
## 80 B for list structure

pryr::object_size(restored_list_of_x)
## 3 * bytes of x + 80 B list structure
```

Using `pickleR` we can serialize and deserialize objects while maintaining the
internal references to common values.

```{r}
library(pickleR)

pickleR_restored_list_of_x <-
  unpickle(pickle(list_of_x, connection = NULL))

pryr::inspect(pickleR_restored_list_of_x)

identical(list_of_x, pickleR_restored_list_of_x)
```

Therefore, using `pickle` and `unpickle` we only have a single copy of `x` in
memory for the `pickle_restored_list_of_x`.

```{r}
pryr::object_size(pickleR_restored_list_of_x)
```

The above example demostrates the `base::` (de-)serialization method does not 
maintain the references to shared objects, but in this case the impact is minimal.
As the objects become increasingly complex or nested, the impact increases. Suppose
we are working with R6 classes, the instances of that object recieve references to the
function bodies from the R6 class generator.

```{r}
library(R6)

classA <-
  R6Class(
    classname = "classA",
    class = TRUE,
    cloneable = TRUE,
    public = list(normalValue = rnorm(1),
                  uniformValue = runif(1),
                  normalRange = function(mean, sd, n = 10) {
                    return(rnorm(n = n, mean = mean, sd = sd))
                  },
                  uniformRange = function(min, max, n = 10) {
                    return(runif(n = n, min = min, max = max))
                  })
  )

instance_of_classA <-
  classA$new()

pryr::object_size(instance_of_classA)

pryr::inspect(instance_of_classA$normalRange)[["children"]][["body"]][["address"]]

second_instance_of_classA <-
  classA$new()

pryr::inspect(second_instance_of_classA$normalRange)[["children"]][["body"]][["address"]]
```

An R6 class instance can be serialized() to a raw vector and restored via unserialize(),
however this serializes the entire instance object including function bodies, creating new
copies of the function bodies for each reference.

```{r}
restored_instance_of_classA <-
  unserialize(serialize(instance_of_classA, connection = NULL))


pryr::object_size(restored_instance_of_classA)

pryr::inspect(restored_instance_of_classA$normalRange)[["children"]][["body"]][["address"]]

```

Using pickleR we serialize and unserialize the state of the instance, not the
instance itself. 

```{r}

pickleR_restored_instance_of_classA <-
    unpickle(pickle(instance_of_classA, connection = NULL))

pryr::object_size(pickleR_restored_instance_of_classA)

pryr::inspect(pickleR_restored_instance_of_classA$normalRange)[["children"]][["body"]][["address"]]
```

The 'pickle' information contains the state of the instance, its class and the
namespace where the class generator is defined. The 'unpickle' function,
instantiates a new instance of the R6 class bypassing the `initialize()` call.
Then uses the pickled state to set the instance fields. The restored instance
recieve references to the function bodies from the R6 class generator, maintaining
internal references and producing an instance of equal memory size to the original.

This example can be extended with a list of R6 instances:

```{r}
list_of_classA_instances <-
  lapply(X = 1:10,
         FUN = function(i) {
           classA$new()
         })

pryr::object_size(list_of_classA_instances)
```

Using the `base::` (de-)serialization method results in a list of objects each with
its own function body definition.

```{r}
restored_list_of_classA_instances <-
  unserialize(serialize(list_of_classA_instances, connection = NULL))

pryr::object_size(restored_list_of_classA_instances)

```

But using `pickleR`:

```{r}
pickleR_restored_list_of_classA_instances <-
  unpickle(pickle(list_of_classA_instances, connection = NULL))

pryr::object_size(pickleR_restored_list_of_classA_instances)


```
